The resurgence of the linear optics interferometer --- recent advances \& applications

### 2017-05-14 paper.tex
Revised subsections 6.3 and 9.3.

GHZ states form a class of entangled quantum states on multiple subsystems with at least three parties \cite{bib:GHZ89}. For qubit encodings with $n$ subsystems, a GHZ state is of the form
\begin{align}
\ket{\Psi_{n}^{\rm (GHZ)}}=\frac{\ket{0}^{\otimes n}+\ket{1}^{\otimes n}}{\sqrt{2}} \ .
\end{align}
These states are also non-local, and have been used extensively in experimental test for non-locality \cite{bib:JW00,bib:Zhang15}. This has been a subject matter covered in detail by another review \cite{bib:JW12}.

@@

\subsection{Quantum metrology}
An archetypal task for quantum metrology is the following: Given a Mach-Zedner interferometer, with an unknown phase shift of magnitude $\varphi$ inserted in one of the two paths (see Fig. \ref{fig:MZ}), can one deduce $\varphi$ via a judicious use of quantum states and measurements with a higher sensitivity than that of a completely classical interferometer? It is well-known that one indeed can, and the standard deviation, $\Delta\varphi$, after $N$ trials using quantum estimation is
\begin{align}
\Delta\varphi=\frac{1}{N} \ ,	
\end{align}
which gets a $\sqrt{N}$ improvement over classical phase estimation. As this sensitivity is also a fundamental limit arising from Heisenberg's uncertainty principle; we cannot measure definitively the phase, it is known as the Heisenberg limit. Most famously, a family of entangled states known as NOON states achieve this limit using $N$ photons in a single trial. A NOON state is a superposition state in which $N$ photons are either in the first or second path:
\begin{align}
\ket{\psi^{\rm (NOON)}_n}=\frac{1}{\sqrt{2}}(\ket{N,0}+\ket{0,N})\ .	
\end{align}
 Many reviews have been written for quantum metrology and Heisenberg-limited quantum phase estimation \cite{bib:Giovannetti04,bib:Dowling08,bib:Giovannetti11} so we will not dwell on the history of this application. Instead, we will discuss a new paradigm for quantum metrology that has arisen from our understanding of Boson-sampling.

>>>>

GHZ states form a class of entangled quantum states on multiple subsystems with at least three parties \cite{bib:GHZ89}. For qubit encodings with $n$ subsystems, a GHZ state is of the form
\begin{align}
\ket{\Psi_{n}^{\rm (GHZ)}}=\frac{\ket{0}^{\otimes n}+\ket{1}^{\otimes n}}{\sqrt{2}} \ .
\end{align}
These states are also non-local, and have been used extensively in experimental test for non-locality \cite{bib:JW00,bib:Zhang15}. This has been a subject matter covered in detail by another review \cite{bib:JW12}. To date, six- \cite{bib:Lu06,bib:Zhang15}, eight- \cite{bib:Huang11,bib:Yao12}, and ten-photons \cite{bib:WangChen16,bib:Chen17} GHZ states have been produced.

The GHZ state is also known as the NOON state, because it is written as
\begin{align}
\ket{\psi^{\rm (NOON)}_n}=\frac{1}{\sqrt{2}}(\ket{N,0}+\ket{0,N})\ ,
\end{align}
in the second quantization representation. This is the notation adopted by the community working in quantum metrology (see Section \ref{subsec:metrology}).

@@

\subsection{Quantum metrology}\label{subsec:metrology}
An archetypal task for quantum metrology is the following: Given a Mach-Zedner interferometer, with an unknown phase shift of magnitude $\varphi$ inserted in one of the two paths (see Fig. \ref{fig:MZ}), can one deduce $\varphi$ via a judicious use of quantum states and measurements with a higher sensitivity than that of a completely classical interferometer? Indeed, one can, and the standard deviation, $\Delta\varphi$, after $N$ trials using quantum estimation is $\Delta\varphi=1/N$, which gets a $\sqrt{N}$ improvement over classical phase estimation. As this sensitivity is also a fundamental limit arising from Heisenberg's uncertainty principle; we cannot measure definitively the phase, it is known as the Heisenberg limit. Most famously, NOON states achieve this limit using $N$ photons in a single trial. Many reviews have been written for quantum metrology and Heisenberg-limited quantum phase estimation \cite{bib:Giovannetti04,bib:Dowling08,bib:Giovannetti11} so we will not dwell on the history of this application. Instead, we will discuss a new paradigm for quantum metrology that has arisen from our understanding of Boson-sampling.
